


```{r include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  fig.align = "center",
  # width = 100,
  warning = FALSE,
  message = FALSE
)


# Need this if render script is not in same directory as project root.
# knitr::opts_knit$set(root.dir = "O:/_other/projects/uil/")
knitr::opts_knit$set(root.dir = rprojroot::find_root("uil.Rproj"))

options(scipen = 1, digits = 2)

# Comment these lines out if calling this script with parameters.
# rm(list = ls())
```




```{r include = FALSE}
load("R/analyze_00.RData")
```





```{r include = FALSE}
# Analyze. ----
```



# Analysis

With this knowledge in mind, let's take begin with "basic" analysis, and
work our way towards nddressing more specific questions.



```{r include = FALSE}
# analysis_complvl ----
```


## "Competition Level" Analysis

### Which _districts, regions, and conferences_ have the most _distinct schools_? [^fn_competition_groupings]

[^fn_competition_groupings]:
As a technical note, districts, regions, and conferences are not all of the same "type"
in the data. Rather,
`District` and `Region` are classified as disticnt factors in the `complvl`
column (along with `State`) and conferences are
classified in their own `conf` column. Nevertheless, for the purpose of exploration,
these diferent "competition groupings" each stratify the sample population in some manner.


```{r results = "hide"}
rank_and_arrange <-
  function(d,
           colname_val_char = "cnt",
           colname_rank_char = "rank") {
    colname_val_quo <- rlang::sym(colname_val_char)
    colname_rank_quo <- rlang::sym(colname_rank_char)
    d %>%
      mutate(!!colname_rank_quo := row_number(desc(!!colname_val_quo))) %>%
      arrange(!!colname_rank_quo) %>%
      select(!!colname_rank_quo, everything())
  }

compute_cnt_bycomplvl <- function(complvls = complvls_valid) {
  schools_all %>%
    filter(complvl %in% complvls) %>%
    group_by(year, complvl_num) %>%
    distinct(year, complvl_num, school) %>%
    summarise(cnt = n()) %>%
    ungroup() %>%
    group_by(complvl_num) %>%
    summarise(cnt = mean(cnt)) %>%
    ungroup() %>%
    rank_and_arrange()
  # mutate(rank = row_number(desc(cnt))) %>%
  # arrange(rank)
}
cnt_bycomplvl_district <- compute_cnt_bycomplvl("District")
cnt_bycomplvl_region <- compute_cnt_bycomplvl("Region")

cnt_bycomplvl_byconf <-
  schools_all %>%
  distinct(year, conf, school) %>%
  group_by(year, conf) %>%
  summarise(cnt = n()) %>%
  ungroup() %>%
  group_by(conf) %>%
  summarise(cnt = mean(cnt)) %>%
  ungroup() %>%
  rank_and_arrange()
```





```{r fig.show = "hide"}
labs_cnt_byx <-
  labs(x = NULL, y = NULL)
theme_cnt_byx <-
  theme(# panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  theme(legend.position = "none")

visualize_cnt_byx <-
  function(d,
           x_char = "complvl_num",
           subtitle_suffix = "") {
    d %>%
      ggplot(aes_string(x = x_char, y = "cnt")) +
      # Coluld make color an optional input.
      geom_point(aes(size = desc(rank)), color = color_neutral_1) +
      # scale_color_distiller(palette = "Reds") +
      # hrbrthemes::scale_color_ipsum() +
      scale_y_continuous(breaks = scales::pretty_breaks()) +
      labs(title = "Count of Schools",
           subtitle = str_c("By ", subtitle_suffix)) +
      labs_cnt_byx +
      theme_cnt_byx
  }

viz_cnt_bycomplvl_district <-
  cnt_bycomplvl_district %>%
  visualize_cnt_byx(x_char = "complvl_num", subtitle_suffix = "District")
viz_cnt_bycomplvl_district

viz_cnt_bycomplvl_region <-
  cnt_bycomplvl_region %>%
  visualize_cnt_byx(x_char = "complvl_num", subtitle_suffix = "Region")
viz_cnt_bycomplvl_region

viz_cnt_bycomplvl_byconf <-
  cnt_bycomplvl_byconf %>%
  visualize_cnt_byx(x_char = "conf", subtitle_suffix = "Conference")
viz_cnt_bycomplvl_byconf
```





```{r include = FALSE}
# {viz_cnt_bycomplvl_district + viz_cnt_bycomplvl_region} / viz_cnt_bycomplvl_byconf + plot_layout(ncol = 1)
```





```{r }
# bookmark ----
viz_cnt_bycomplvl_grid <-
  arrangeGrob(
    viz_cnt_bycomplvl_district,
    viz_cnt_bycomplvl_region + labs(title = NULL),
    viz_cnt_bycomplvl_byconf + labs(title = NULL),
    ncol = 2,
    nrow = 2,
    # top = textGrob("Count of Schools", gp = gpar(fontsize = 14, family = "text", font = 2),
    layout_matrix = rbind(c(1, 1), c(2, 3))
  )
grid.arrange(viz_cnt_bycomplvl_grid)
save_viz(viz_cnt_bycomplvl_grid)
```


It seems fair to say that the distribution of schools among districts/regions/conferences
is relatively even. This is to be expected since the UIL (presumably)
tries to divide schools evenly among each grouping (to the extent possible) in
order to stimulate fair competition.

### Which competition types have the most individual competitors and distinct schools?


```{r results = "hide"}
compute_cnt_bycomp_byx <- function(d) {
  d %>%
    group_by(year, comp) %>%
    summarise(cnt = n()) %>%
    ungroup() %>%
    group_by(comp) %>%
    summarise(cnt = mean(cnt)) %>%
    ungroup() %>%
    rank_and_arrange()
}
cnt_bycomp_byperson <- persons_all %>% compute_cnt_bycomp_byx()
cnt_bycomp_byschool <- schools_all %>% compute_cnt_bycomp_byx()
```





```{r fig.show = "hide"}
lab_persons <- "Individual Competitors"
lab_schools <- "Distinct Schools"
visualize_cnt_bycomp_byx_base <-
  function(d,
           x_char = "comp",
           color_char = x_char,
           title_suffix = "",
           lab_subtitle = "By Competition Type") {
    d %>%
      ggplot(aes_string(x = x_char, y = "cnt")) +
      geom_point(aes_string(color = color_char), size = 4) +
      # geom_point(aes(color = comp, size = desc(rank))) +
      # scale_radius(range = c(3, 7)) +
      scale_x_discrete(labels = scales::wrap_format(width_xy_wrap)) +
      scale_y_continuous(breaks = scales::pretty_breaks()) +
      labs(title = str_c("Count of ", title_suffix),
           subtitle = lab_subtitle) +
      labs_cnt_byx +
      theme_cnt_byx
  }


visualize_cnt_bycomp_byx <- function(d, title_suffix = "") {
  d %>%
    visualize_cnt_bycomp_byx_base(x_char = "comp",
                                  title_suffix = title_suffix,
                                  lab_subtitle = "By Competition Type")

}

viz_cnt_bycomp_byperson <-
  cnt_bycomp_byperson %>%
  visualize_cnt_bycomp_byx(title_suffix = lab_persons)
viz_cnt_bycomp_byperson

viz_cnt_bycomp_byschool <-
  cnt_bycomp_byschool %>%
  visualize_cnt_bycomp_byx(title_suffix = lab_schools)
viz_cnt_bycomp_byschool
```





```{r }
viz_cnt_comp_byx_grid <-
  arrangeGrob(
  viz_cnt_bycomp_byperson + labs(subtitle = NULL),
  viz_cnt_bycomp_byschool + labs(subtitle = NULL),
  nrow = 2
)
grid.arrange(viz_cnt_comp_byx_grid)
save_viz(viz_cnt_comp_byx_grid)
```


####  ... at _each competition level_?


```{r results = "hide"}
compute_cnt_bycomp_byx_bycomplvl <- function(d) {
  d %>%
    group_by(year, complvl, comp) %>%
    summarise(cnt = n()) %>%
    ungroup() %>%
    group_by(complvl, comp) %>%
    summarise(cnt = mean(cnt)) %>%
    ungroup() %>%
    group_by(complvl) %>%
    mutate(rank = row_number(desc(cnt))) %>%
    ungroup() %>%
    arrange(complvl, rank)
}
cnt_bycomp_byperson_bycomplvl <-
  persons_all %>% compute_cnt_bycomp_byx_bycomplvl()
cnt_bycomp_byschool_bycomplvl <-
  schools_all %>% compute_cnt_bycomp_byx_bycomplvl()
```





```{r fig.show = "hide"}
visualize_cnt_bycomp_byx_bycomplvl <-
  function(d, title_suffix = "") {
    d %>%
      visualize_cnt_bycomp_byx_base(x_char = "comp",
                                    title_suffix = title_suffix,
                                    lab_subtitle = "By Competition Type and Competition Level") +
      facet_grid(complvl ~ .,
                 scales = "free",
                 labeller = label_wrap_gen(width = width_xy_wrap))
  }
viz_cnt_bycomp_byperson_bycomplvl <-
  cnt_bycomp_byperson_bycomplvl %>%
  visualize_cnt_bycomp_byx_bycomplvl(title_suffix = lab_persons)
viz_cnt_bycomp_byperson_bycomplvl

viz_cnt_bycomp_byschool_bycomplvl <-
  cnt_bycomp_byschool_bycomplvl %>%
  visualize_cnt_bycomp_byx_bycomplvl(title_suffix = lab_schools)
viz_cnt_bycomp_byschool_bycomplvl
```





```{r }
viz_cnt_bycomp_byx_bycomplvl_grid <-
  arrangeGrob(
    viz_cnt_bycomp_byperson_bycomplvl + labs(subtitle = NULL),
    viz_cnt_bycomp_byschool_bycomplvl + labs(subtitle = NULL),
    ncol = 2
  )
grid.arrange(viz_cnt_bycomp_byx_bycomplvl_grid)
save_viz(viz_cnt_bycomp_byx_bycomplvl_grid)
```


#### ... in _each conference_?


```{r results = "hide"}
compute_cnt_bycomp_byx_bycomplvl_byconf <- function(d) {
  d %>%
    group_by(year, complvl, comp, conf) %>%
    summarise(cnt = n()) %>%
    ungroup() %>%
    group_by(complvl, comp, conf) %>%
    summarise(cnt = mean(cnt)) %>%
    ungroup() %>%
    group_by(complvl, comp) %>%
    mutate(rank = row_number(desc(cnt))) %>%
    ungroup() %>%
    arrange(comp, complvl, rank)
}

cnt_bycomp_byperson_bycomplvl_byconf <-
  persons_all %>%
  compute_cnt_bycomp_byx_bycomplvl_byconf()
cnt_bycomp_byschool_bycomplvl_byconf <-
  schools_all %>%
  compute_cnt_bycomp_byx_bycomplvl_byconf()
```





```{r }
visualize_cnt_bycomp_byx_bycomplvl_byconf <-
  function(d, title_suffix = "") {
    d %>%
      visualize_cnt_bycomp_byx_base(
        x_char = "conf",
        color_char = "comp",
        title_suffix = title_suffix,
        lab_subtitle = "By Competition Type, Competition Level, and Conference"
      ) +
      facet_grid(complvl ~ comp,
                 scales = "free",
                 labeller = label_wrap_gen(width = width_xy_wrap))
    # facet_grid(comp ~ complvl, scales = "free", labeller = label_wrap_gen(width = width_xy_wrap))
  }
viz_cnt_bycomp_byperson_bycomplvl_byconf <-
  cnt_bycomp_byperson_bycomplvl_byconf %>%
  visualize_cnt_bycomp_byx_bycomplvl_byconf(title_suffix = lab_persons)
viz_cnt_bycomp_byperson_bycomplvl_byconf
save_viz(viz_cnt_bycomp_byperson_bycomplvl_byconf)

viz_cnt_bycomp_byschool_bycomplvl_byconf <-
  cnt_bycomp_byschool_bycomplvl_byconf %>%
  visualize_cnt_bycomp_byx_bycomplvl_byconf(title_suffix = lab_schools)
viz_cnt_bycomp_byschool_bycomplvl_byconf
save_viz(viz_cnt_bycomp_byschool_bycomplvl_byconf)
```


Just by inspection, science appears to be the answer to the numerous
variations of the "Which competition type/level has the most ...?" question.



```{r include = FALSE}
# # Attempt to make a NSE version that combines the functionality of the previous two functions.
# compute_cnt_bycomp_byx_nse <- function(d, complvls = complvls_valid, ...){
#   group_1_vars_quos <- rlang::quos(...)
#   # group_1_vars_quos <- quos(group_1_vars_quos)
#   # group_1_vars_quos <- rlang::quos(year, complvl, comp_shortname, comp, conf)
#
#   group_2_vars_quos <- setdiff(rlang::quo_name(group_1_vars_quos), "year")
#   # group_2_vars_quos <- rlang::quos(complvl, comp_shortname, comp, conf)
#   group_3_vars_quos <- rlang::quos(complvl, conf)
#   d %>%
#     filter(complvl %in% complvls) %>%
#     group_by(!!!group_1_vars_quos) %>%
#     summarise(cnt = n()) %>%
#     ungroup() %>%
#     group_by(!!!group_2_vars_quos) %>%
#     summarise(cnt = mean(cnt)) %>%
#     ungroup() %>%
#     group_by(!!!group_3_vars_quos) %>%
#     mutate(rank = row_number(desc(cnt))) %>%
#     arrange(rank)
# }
```






```{r include = FALSE}
session::save.session("R/analyze_01.RData")
```




